
---
title: "Mini-Project #01: What We Watched"
author: "Rojina Adhikari"
format:
  html:
    code-fold: true
---
  
## Introduction
  
  Netflix publishes its *What We Watched (WWW)* report every six months, beginning in 2023. This report analyzes  *Netflix's What We Watched* data to explore patterns in viewership data, identify the type of content, understand which content captures the most attention and who are the audiences in order to make informed decision making.
  
  The goal is not just to analyze and provide  practical recommendations for Netflix's 2027 content budget but also to help Netflix unlock broader patterns to make more informed decisions about their future investment decisions. 
  
## Data Acquisition
  
```{r}
#| label: load-data
#| message: false
#| warning: false
load_netflix_data <- function(){

  library(tidyverse)
  library(rvest)
  library(readxl)
  library(fs)
  
  data_dir <- path("data", "mp01")
    
  if(!dir_exists(data_dir)){
    dir_create(data_dir, showWarnings=FALSE, recurse=TRUE)
  }
  
  NETFLIX_URLS <- list(
    `2023H1` = "https://about.netflix.com/news/what-we-watched-a-netflix-engagement-report",
    `2023H2` = "https://about.netflix.com/news/what-we-watched-the-second-half-of-2023",
    `2024H1` = "https://about.netflix.com/news/what-we-watched-the-first-half-of-2024",
    `2024H2` = "https://about.netflix.com/news/what-we-watched-the-second-half-of-2024",
    `2025H1` = "https://about.netflix.com/news/what-we-watched-the-first-half-of-2025",
    `2025H2` = "https://about.netflix.com/news/what-we-watched-the-second-half-of-2025",
    `2026H1` = "https://about.netflix.com/news/what-we-watched-the-first-half-of-2026"
  )
  
  download_if_needed <- function(url, local){
    if(!file_exists(local)){
      download.file(url, local, quiet = TRUE, mode = "wb")
    }
    
    return(local)
  }
    
  NETFLIX_URLS |> 
    map(read_html) |> 
    map(html_element, xpath = "//a[contains(@href, '.xlsx')]") |> 
    map_chr(html_attr, "href") |> 
    enframe("period", "url") |> 
   mutate(path = path(data_dir, paste0(period, ".xlsx")), 
          path = map2(url, path, download_if_needed), 
          data = map(path, read_excel, skip=3, na="*"), 
          data = map(data, \(x) x |> mutate(`Release Date` = as.Date(`Release Date`)))) |>
    unnest(data) |> 
    select(-url, -path) |> 
    mutate(year=str_sub(period, 1, 4), 
           half=str_sub(period, start=6), 
           period_start = make_date(year, if_else(half==1, 1, 7), 1), 
           period_end = make_date(year, if_else(half==1, 7, 12), if_else(half==1, 30, 31))) |> 
    filter(!is.na(`Hours Viewed`)) |> 
    mutate(live_event = !is.na(Runtime) & str_detect(Runtime, "\\d+:\\d+\\*"), 
           clean_time = str_remove(Runtime, fixed("*")),
           Runtime = 60 * hour(hm(clean_time)) + minute(hm(clean_time))) |>
    select(-clean_time) |>
    mutate(Runtime = coalesce(Runtime, mean(Runtime, na.rm=TRUE)), 
           .by=Title, 
           Title = stringi::stri_trans_general(Title, "Latin-ASCII"),
           Title = str_remove(Title, " // .*"), 
           Title = str_remove(Title, "[^[:ascii:]].*"), 
           Title = str_trim(Title)) |>
    filter(Title != "") |>
    select(year, 
           period, 
           period_start, 
           period_end, 
           Title, 
           -half, 
           Runtime, 
           Views, 
           `Hours Viewed`, 
           `Release Date`,
           `Available Globally?`, 
           live_event) |>
    rename(                 ## Task 2: Fix Column Titles
          title = Title,
          runtime = Runtime,
          views = Views,
          hours_viewed = `Hours Viewed`,
          release_date = `Release Date`,
          available_globally = `Available Globally?`
          )|>                                           # <- pipe added and task 3:Fix column types
    mutate(
      year = as.integer(year),
      available_globally = available_globally == "Yes"
    )
  }
  
  NETFLIX_DATA <- load_netflix_data()

  
```
  
## Data Cleaning and preparation
 Before analysis, the raw data needed two fixes: messy column names 
and incorrect data types. Both were added directly to the 
`load_netflix_data()` function above — column names were converted to 
lowercase with underscores (e.g. `Hours Viewed` → `hours_viewed`), 
`year` was converted to an integer, and `available_globally` was 
converted from `"Yes"`/`"No"` text to `TRUE`/`FALSE` logical values.

```{r}
#| label: verify-cleaning

glimpse(NETFLIX_DATA)
```
 
## Exploratory Analysis

### Initial Data Exploration  
 Before answering any questions, I ran two basic checks on the cleaned
data: one for column types and one for summary statistics.

#### Check 1: Column types

```{r}
#| label: Data_Exploration_glimpse

glimpse(NETFLIX_DATA)
```
[2-3 sentences: Are the types what you expect? Mention that `year` is
now an integer, `available_globally` and `live_event` are logical, and
the three date columns are dates.]

#### Check 2: Summary Statistics

```{r}
#| label: Data_Exploration_skim
 library(skimr)
 skim(NETFLIX_DATA)
``` 
[3-5 sentences: What stands out? Cover:
- how many rows, and the date range covered
- which columns have missing values (`views`, `runtime`) and roughly how many
- how skewed `hours_viewed` is (compare the median to the max)
- what this means for the analysis, for example that `views` can't be
  relied on because it's mostly missing]
  
  Tips for the description:

Write about what the numbers imply for your analysis, not just what they are. "views is missing for most titles, so I use hours_viewed as the main measure" shows more thinking than "views has missing values."
Take the numbers from your skim() output, such as n_missing and p50 versus p100.
Mention the data limitations the assignment itself notes, such as hours being rounded to the nearest 100,000.


### Task 5- Exploratory questions

[1-2 sentences of your own: what these six questions are for, e.g. getting
a sense of the data's scale and where viewing is concentrated.]

```{r}
#| label: task5-values

# Q1: date range (period_end is off for H1 periods, so derive the end date)
first_date <- min(NETFLIX_DATA$period_start)
last_date  <- max(NETFLIX_DATA$period_start) + months(6) - days(1)

# Q2: number of distinct shows and movies (seasons count separately)
n_titles <- n_distinct(NETFLIX_DATA$title)

# Q3: total hours for the Bridgerton universe
bridgerton_hours <- NETFLIX_DATA |>
  filter(str_detect(title, "Bridgerton")) |>
  summarise(total = sum(hours_viewed)) |>
  pull(total)

# Q4: total hours for the five Stranger Things seasons
stranger_hours <- NETFLIX_DATA |>
  filter(str_detect(title, "Stranger Things")) |>
  filter_out(str_detect(title, "Beyond Stranger Things")) |>
  filter_out(str_detect(title, regex("Tales from", ignore_case = TRUE))) |>
  filter_out(str_detect(title, "Fireplace")) |>
  summarise(total = sum(hours_viewed)) |>
  pull(total)

# Q5: share of 2026H1 watch time spent on that period's ten biggest titles
h1_2026 <- NETFLIX_DATA |> filter(period == "2026H1")
top10_share <- h1_2026 |>
  slice_max(hours_viewed, n = 10, with_ties = FALSE) |>
  summarise(top10 = sum(hours_viewed)) |>
  pull(top10) / sum(h1_2026$hours_viewed)

# Q6: share of catalog that is globally available
# Q6: share of the catalog that is globally available
# By title: every title-period listing counts once
global_by_title <- mean(NETFLIX_DATA$available_globally, na.rm = TRUE)

# Weighted by hours viewed
global_by_hours <- weighted.mean(
  NETFLIX_DATA$available_globally,
  NETFLIX_DATA$hours_viewed,
  na.rm = TRUE
)
```


### Exploratory Questions-Inline Values

[Your 1-2 sentences: what these six questions check and why.]

(chunk above goes here)

[Paragraph 1, Q1 and Q2: the period covered and the number of titles.
Inline code: `r format(first_date, "%B %d, %Y")`,
`r format(last_date, "%B %d, %Y")`, `r scales::comma(n_titles)`.
Note that seasons are counted separately.]

[Paragraph 2, Q3 and Q4: hours for each franchise.
Inline code: `r scales::comma(bridgerton_hours)`,
`r scales::comma(stranger_hours)`.
State what you included in Bridgerton and which three Stranger Things
companion titles you excluded.]

[Paragraph 3, Q5 and Q6: concentration and availability.
Inline code: `r scales::percent(top10_share, accuracy = 0.1)`,
`r scales::percent(global_by_title, accuracy = 0.1)`,
`r scales::percent(global_by_hours, accuracy = 0.1)`.
State how you read "top 10", and say what the gap between the two
availability numbers means.]

[Your 1-2 closing sentences tying the answers together.]

### Task 6: Exploratory Questions - Table Answers

```{r}
#| label: load-gt
#| message: false

library(gt)
```
#### 1. What was the top show (by total hours viewed) in each period and what fraction of total viewership did it capture?

```{r}
#| label: task6-catalog
top_by_period <- NETFLIX_DATA |>
  mutate(period_total = sum(hours_viewed), .by = period) |>
  slice_max(hours_viewed, n = 1, by = period, with_ties = FALSE) |>
  mutate(share = hours_viewed / period_total) |>
  select(period, title, hours_viewed, share) |>
  arrange(period)

top_by_period |>
  gt() |>
  cols_label(period = "Period", title = "Top Title",
             hours_viewed = "Hours Viewed", share = "Share of Period Viewing") |>
  fmt_number(hours_viewed, decimals = 1, suffixing = TRUE) |>
  fmt_percent(share, decimals = 1) |>
  tab_header(title = "[Your one-line takeaway]") |>
  tab_source_note("Source: Netflix What We Watched reports")
```

  
#### State: "Show" means one title row, so each season counts separately. The share is out of the hours that Netflix reports for that period, since low-viewership titles are omitted.

#### How has Netflix’s catalog changed in size over the periods for which we have questions? You should consider both the number of titles and the total runtime of content.

```{r}
#| label: task6-catalog
catalog_size <- NETFLIX_DATA |>
  summarise(n_titles = n(),
            total_runtime_hours = sum(runtime, na.rm = TRUE) / 60,
            .by = period) |>
  arrange(period)

catalog_size |>
  gt() |>
  cols_label(period = "Period", n_titles = "Titles Reported",
             total_runtime_hours = "Total Runtime (hours)") |>
  fmt_number(c(n_titles, total_runtime_hours), decimals = 0) |>
  tab_header(title = "[Your one-line takeaway]") |>
  tab_source_note("Source: Netflix What We Watched reports")
```

#### State: This counts titles that appear in the reports, not Netflix's full library, because titles under the viewing threshold are left out. Runtime is missing for some titles and is excluded from the sum.

#### Which shows have had the most consistently high viewership across all periods? These shows are likely to be ‘staples’ that viewers fall back on as ‘comfort food’, e.g., re-runs of Friends.

#### Within each period, what fraction of time was spent on ‘shows’ vs movies?

#### How does the performance of live events compare to pre-taped content? Does the average live event draw more viewership than an average piece of pre-taped content? How about a ‘top tier’ live event vs ‘top tier’ pre-taped content?
